Improving content discovery through combining linked data and data mining techniques

### 2016-05-06 report/chapters/01-intro.tex
Fixed broken references

Many of these problems are felt within the BBC and are potentially
issues in many similar organisations. The BBC website officially
launched in 1997 and many articles from that year are still available
in archived form. Over the last 20 years, multiple content management
systems were created and retired such that today even the active
website is served by a mix of emerging, cutting-edge platforms and
systems that have been running for a number of years.

Technologies such as the World Wide Web and the maturity of the
media industry in its adoption in the last decade creates both an
opportunity to give consumers access to a substantial amount of
media content and a challenge in trying to make cohesive sense of
a diverse range of content, perhaps produced in an organisation that
was traditionally very fragmented as the BBC was.

@@

It is proposed also that all the data acquisition, enrichment and
mappings can be implemented in a general way when semantics and
linked data are used. The standardisations offered by the Semantic
Web and Linked Data research fields provides uniform interfaces and
patterns of data manipulation such that any software produced can be
applied to all of an organisation's content without consideration to
any particular class or type of item. This also does not preclude
including custom business rules later either, where domain experts
can give more guidance to treat data more intelligently.

>>>>

Many of these problems are felt within the BBC and are potentially
issues in many similar organisations. The BBC website officially
launched in 1997 and many articles even from that year are still available
in archived form. Over the last 20 years, multiple content management
systems were created and retired such that today even the active
website is served by a mix of emerging, cutting-edge platforms and
systems that have been running for a number of years.

@@

It is proposed also that all the data acquisition, enrichment and
mappings can be implemented in a general way when semantics and
linked data are used. The standardisations offered by the Semantic
Web and Linked Data research fields provide uniform interfaces and
patterns of data manipulation such that any software produced can be
applied to all of an organisation's content without consideration to
any particular class or type of item. This also does not preclude
including custom business rules later either, where domain experts
can give more guidance to treat data more intelligently.

### 2016-05-06 report/chapters/02-background.tex
Fixed broken references

\begin{enumerate}
\item that extracting feature sets from
  RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done
  elegantly and follows naturally from some previous work in this area; and
\item that the RDF graph is a suitable and even desirable data model for content
  metadata in terms of acquiring, enriching and even transforming that data ahead
  of feature extraction.
\end{enumerate}

@@

\emph{Data Mining} is a process for finding knowledge and information
in data sets by drawing on research from artificial intelligence,
machine learning and statistics.\cite{han2011data} In this research,
the focus is on the machine learning aspect with
the data extraction phase implemented using semantics and RDF.

@@

Well-known examples include neural networks, decision trees and or
linear regression. In all supervised learning techniques, we can
normally assess the model's performance against training data for
which we already know the answers. We can see supervised learning
thus as a computational attempt to extrapolate human knowledge into
a reusable model that can predict or determine things at a speed
and scale humans would not normally be able to work through manually
(e.g. predicting weather with manual calculations or diagnosing
patients more quickly and with greater accuracy).

@@

In unsupervised learning, there is not necessarily a specific goal
against which we can judge the models, but instead we are applying
more exploratory methods to tell us patterns in the data that maybe
we could not see for ourselves (e.g. if we would have to graph trends
over more than two or three dimensions, it is usually difficult to
visualise).

@@

Another major unsupervised learning technique -- and the focus of
this research -- is \emph{clustering}. In clustering algorithm, we
seek to group similar items or entities together in groups or
\emph{clusters}. Clustering has many applications, but in this
work the focus will be around \emph{hierarchical} clustering which
is effective for grouping together items that appear to be
connected in some way -- in our case that the media content is
connected by virtue of a common theme, topic or genre.

@@

Clustering relies on grouping items that seem close, which creates
a necessity to define what is meant by close. The distance between
any two items is normally determined by a distance
function.\cite{witten2005data} Such a function will normally take
two items and return a value between 0 (items are the same) up to an
an unbounded value as the items are considering further and further
apart.

@@

In this experiment, Jaccard distance\cite{witten2005data}
based on its suitability for binary, asymmetric data. It is argued
that the vast vocabulary offered by all potential semantic web
RDF triples will lead to very asymmetric binary values based on the
presence or non-presence of that triple.

@@

Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
approach called SPARQL-ML -- an extension to the
SPARQL\cite{segaran2009programming} query language with new keywords to
facilitate both generating and applying models. This means that the system
capable of parsing and running standard queries must also run machine learning
algorithms.

@@

@@

\noindent this query could then be
expressed as function $f: G \rightarrow (S \times S)$ where $G$ is the set of
all possible RDF graphs and $S$ is a set of all possible strings.
This allows the result of the
SPARQL query to be expressed as a set of all SELECT variable bindings that
satisfy the WHERE clause:

@@

This might allow a query that extracts a country's population, GDP, etc.
provide feature extraction for learning patterns in economics, for example.
However, this is limited to features derived from single-valued predicates
with literal-valued ranges. It is not clear how to formulate a query that
expresses whether or not a content item is about a given topic.

@@

@@

In this section, the goal is to strengthen that case and review other
previous work around using linked data in the enterprise. This is a more
general argument for any enterprise with data (e.g. internal information
about staff or business processes) and not just media organisations.

@@

It is rare, however, for any large business to keep all data stored in
one location and in one format. Even fairly simple businesses will have
separate data stores for their employee data, for example, and all
user names and passwords for the corporate email system. Without much effort,
a business that implements or purchases two distinct systems for personnel
and email has already fragmented its data.

>>>>

\begin{enumerate}
\item that extracting feature sets from
  RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done
  elegantly and follows naturally from some previous work in this area;
\item that the RDF graph is a suitable and even desirable data model for content
  metadata in terms of acquiring, enriching and even transforming that data ahead
  of feature extraction;
  \item the use of semantics and RDF models has many benefits in general
  for an organisation needing to reason across all of its data.
\end{enumerate}

@@

\emph{Data Mining} is a process for finding knowledge and information
in data sets by drawing on research from artificial intelligence,
machine learning and statistics.\cite{han2011data} In this research,
the focus is on the machine learning aspect with
the feature extraction phase implemented using semantics and RDF.

@@

Well-known examples include neural networks, decision trees and or
linear regression. In all supervised learning techniques, we can
normally assess the model's performance against training data for
which we already know the answers. We can see supervised learning
thus as a computational attempt to extrapolate human knowledge into
a reusable model that can predict or determine things at a speed
and scale humans would not normally be able to work through manually
(e.g. predicting weather without manual calculations or diagnosing
patients more quickly and with greater accuracy).

@@

In unsupervised learning, there is not necessarily a specific goal
against which we can judge the models, but instead we are applying
more exploratory methods to tell us patterns in the data that maybe
we could not easily see for ourselves. If we would have to graph trends
over more than two or three dimensions, it is usually difficult to
visualise. When we considered items using a tag-based taxonomy to be
multi-dimensional vectors of binary attributes (presence or non-presence
of a tag), we easily open up the need to reason across thousands or even
millions of dimensions.

@@

Another major unsupervised learning technique -- and the focus of
this research -- is \emph{clustering}. In clustering algorithm, we
seek to group similar items or entities together into groups or
\emph{clusters}. Clustering has many applications, but in this
work the focus will be around \emph{hierarchical} clustering which
is effective for grouping together items that appear to be
connected in some way -- in our case that the media content is
connected by virtue of a common theme, topic or genre.

@@

Clustering relies on grouping items that seem close, which creates
a necessity to define what is meant by ``close''. The distance between
any two items is normally determined by a distance
function.\cite{witten2005data} Such a function will normally take
two items and return a value between $0$ (items are the same) up to an
an unbounded value as the items are considered further and further
apart.

@@

In this experiment, Jaccard distance\cite{witten2005data} is used
based on its suitability for binary, asymmetric data.\cite{choi2010survey}
It is argued
that the vast vocabulary offered by all potential semantic web
RDF triples will lead to very asymmetric binary values based on the
presence or non-presence of that triple.

@@

Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
approach called SPARQL-ML -- an extension to the
SPARQL\cite{segaran2009programming} query language with new keywords to
facilitate both generating and applying models. However, this means that the system
capable of parsing and running standard queries must also run machine learning
algorithms -- it may be better to decouple these functions in a distributed
system for the ability to tune and scale each independently.

@@

We might be able to think of the set of variable bindings returned from a
SPARQL query as a feature set in their own right. In the example in
listings~\ref{lst:sparqlfoaf}, we can think of the name-email result pair
as feature sets or vectors $(n,e) \in (S \times S)$ where $S$ is the set of
all possible strings.

@@

\noindent this query could then be
expressed as function $f: G \rightarrow (S \times S)$ where $G$ is the set of
all possible RDF graphs.
This allows the result of the
SPARQL query to be expressed as a set of all SELECT variable bindings that
satisfy the WHERE clause:

@@

This might allow a query that extracts a country's population, GDP, etc.
provide feature extraction for learning patterns in economics, for example.
However, this is limited to features derived from single-valued predicates
with literal-valued ranges. It is not immediately clear how to formulate a
query that expresses whether or not a content item is about a given topic.

@@

The use of SPARQL in their work has many similarties to the model proposed
above to create feature sets based on query results, so this research uses
that work as the main inspiration for feature set generation.

@@

In the following sections, the goal is to strengthen that case and review other
previous work around using linked data in the enterprise. This is a more
general argument for any enterprise with data (e.g. internal information
about staff or business processes) and not just media organisations.

@@

It is rare, however, for any large business to keep all data stored in
one location and in one format. For example, even fairly simple businesses will have
separate data stores for their employee data and all
user names and passwords for the corporate email system. Without much effort,
a business that implements or purchases two distinct systems for personnel
and email has already fragmented its data.

### 2016-05-06 report/chapters/03-design.tex
Fixed broken references

\section{Technical Architecture}

@@

        IRI -> RDFa [label="dereference"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="dereference/parse"];
        RDF -> RDF2 [label="dereference/parse"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="content\_?p\_?v"];
      }
    \end{dot2tex}
  \end{center}
  \caption{\label{fig:object.-deref}Semantic web content miner with additional dereferencing of linked entities}
\end{sidewaysfigure}

@@

One of the advantages of the RDF graph model is that we are not constrained
necessarily to properties and attributes directly applicable to the entity. We
could imagine adding a level of indirection to the query in
Listings~\ref{simple-sparql} to create the query in
Listings~\ref{level2-sparql}.

>>>>

\section{High-Level Architecture}

@@

        IRI -> RDFa [label="dereference"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="dereference/parse"];
        RDF -> RDF2 [label="dereference/parse"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="content\_?p\_?v"];
      }
    \end{dot2tex}
  \end{center}
  \caption{\label{fig:object-deref}Semantic web content miner with additional dereferencing of linked entities}
\end{sidewaysfigure}

@@

One of the advantages of the RDF graph model is that we are not constrained
necessarily to properties and attributes directly applicable to the entity. We
could imagine adding a level of indirection to the query in
Listings~\ref{lst:simple-sparql} to create the query in
Listings~\ref{lst:level2-sparql}.

### 2016-05-06 report/chapters/04-implementation.tex
Fixed broken references

For this, there is an
appeal to designing such a system on a cloud computing platform such
that it can be scaled up for the initial import of existing data and
then back down for ongoing updates when that import completes. Such
design is out of scope for this paper, but some discussion of
limitations of the existing system in section~\ref{} highlights where
this system may be architected differently to be more suitable for
such a cloud-based deployment.

@@

An illustrative Python function is shown in
Listings~\ref{lst:extract-all} where each of the three extraction
techniques is invoked only once and a power set function is used to
generate all possible unions of those graphs and thus yield all
possible usages or non-usages of each strategy to generate RDF
graphs. The details of the three extraction functions are covered in
section~\ref{impl-extraction}.

@@

As introduced in section~\ref{intro-problems}, all that content is
also likely to be distributed across multiple databases, content
management systems and other stores. While it is hypothesised in
this research that semantics and linked data provide a good way to
extract data about the total breadth of all content without any
bespoke integration against internal systems, there still remains
the enterprise integration problem of \emph{discovery} of those
content items.

@@

In order to maintain focus in the experiment, only one of the
enrichment techniques described in chapter~\ref{design} was
implemented and its usage and non-usage merely doubled the
number of systems for comparison to fourteen.

@@

\begin{centering}
  \begin{lstlisting}[
      label=lst:enrich,
      language=python,
      caption={Python function that enriches a graph via dereferencing objects}]
import rfc3987

>>>>

For this, there is an
appeal to designing such a system on a cloud computing platform such
that it can be scaled up for the initial import of existing data and
then back down for ongoing updates when that import completes. Such
design is out of scope for this paper, but some discussion in chapter~\ref{chp:evaluation}
of performance limitations of the experiemtanl system highlights where
this system may be architected differently to be more suitable for
such a cloud-based deployment.

@@

An illustrative Python function is shown in
Listings~\ref{lst:extract-all} where each of the three extraction
techniques is invoked only once and a power set function is used to
generate all possible unions of those graphs and thus yield all
possible usages or non-usages of each strategy to generate RDF
graphs. The details of the three extraction functions are covered in
section~\ref{sec:impl-extraction}.

@@

As introduced in section~\ref{sec:intro-problems}, all that content is
also likely to be distributed across multiple databases, content
management systems and other stores. While it is hypothesised in
this research that semantics and linked data provide a good way to
extract data about the total breadth of all content without any
bespoke integration against internal systems, there still remains
the enterprise integration problem of \emph{discovery} of those
content items.

@@

In order to maintain focus in the experiment, only one of the
enrichment techniques described in chapter~\ref{chp:design} was
implemented and its usage and non-usage merely doubled the
number of systems for comparison to fourteen.

@@

\begin{centering}
  \begin{lstlisting}[
      label=lst:enrich,
      showstringspaces=false,
      language=python,
      caption={Python function that enriches a graph via dereferencing objects}]
import rfc3987

### 2016-05-06 report/report.tex
Fixed broken references

However, feedback from the evaluation suggests that real value for
users from suggesting related content will come from linking items
based on topics and themes, which are better found via entity
extraction systems such as DBPedia Spotlight.

>>>>

However, feedback from the evaluation suggests that real value for
users from suggesting related content comes additionally from linking items
based on topics and themes, which are better found via entity
extraction systems such as DBPedia Spotlight.
